---
title: "MA684_homework_08"
author: "Chaoqun Yin"
date: "November 10, 2016"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.width="0.9\\linewidth",dev="pdf",fig.align  = 'center')
library(ggplot2)
library(knitr)
library(gridExtra)
library(arm)
library(data.table)
library(foreign)
library(car)
library(stringr)
library(rstan)
library(zoo)

```


## Getting to know stan
Read through the tutorial on Stan
https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started

* Explore Stan website and Stan reference manual and try to connect them with 
Gelman and Hill 16 - 17.


# Data analysis 

## Using stan:

The folder olympics has seven judges' ratings of seven figure skaters (on two criteria: "technical merit" and "artistic impression") from the 1932 Winter Olympics. Take a look at 
http://www.stat.columbia.edu/~gelman/arm/examples/olympics/olympics1932.txt

```{r,echo=FALSE, fig.width=7,fig.height=3 ,out.width="0.8\\linewidth",message=FALSE}
olympics1932_na<-read.fwf("http://www.stat.columbia.edu/~gelman/arm/examples/olympics/olympics1932.txt",widths=c(2,14,9,9,9,9,9,9,9),skip=21,header = FALSE)
colnames(olympics1932_na)<- c("pair", "criterion", "judge_1",  "judge_2",  "judge_3",  "judge_4",  "judge_5" , "judge_6",  "judge_7")
olympics1932<-na.locf(olympics1932_na)
olympics1932$criterion<-str_trim(olympics1932$criterion)
olympics1932$pair<-str_trim(olympics1932$pair)
ggplot(melt(olympics1932,id.vars=c("pair","criterion")))+geom_point()+aes(x=pair,y=value,group=variable,color=variable)+geom_line()+facet_grid(.~criterion)
molympics<-data.table(melt(olympics1932,id.vars=c("pair","criterion")))
molympics$value <- as.double(molympics$value)
olong <- merge(molympics[seq(1,98,by=2),],molympics[seq(2,98,by=2),],by=c("pair", "variable" ))
setnames(olong,c( "variable",  "value.x", "value.y"),c("Judge","Program","Performance"))
olympics_long <- olong[,list(Program,Performance,pair,Judge)]
head(olympics_long)
pair_country<-str_trim(read.csv("http://www.stat.columbia.edu/~gelman/arm/examples/olympics/olympics1932.txt",skip=3,nrows = 7,header=FALSE,stringsAsFactors=FALSE)$V3)
judge_country<-str_trim(read.csv("http://www.stat.columbia.edu/~gelman/arm/examples/olympics/olympics1932.txt",skip=12,nrows = 7,header=FALSE,stringsAsFactors=FALSE)$V2)
names(pair_country)<-1:7
names(judge_country)<-paste("judge",1:7,sep="_")
olympics_long$same_country<-1*(pair_country[olympics_long$pair]==judge_country[olympics_long$Judge])
```

use stan to fit a non-nested multilevel model (varying across skaters and judges) for the technical merit ratings.

\begin{eqnarray}
y_i &\sim& N(\mu+\gamma_{j[i]}+\delta_{k[i]},\sigma^2_y),\mbox{ for } i=1,\dots, n\\
\gamma_{j} &\sim& N(0,\sigma^2_{\gamma}) j=1,\dots, 7\\
\delta_{k} &\sim& N(0,\sigma^2_{\delta}) k=1,\dots, 7
\end{eqnarray}

https://github.com/stan-dev/example-models/blob/master/ARM/Ch.17/17.3_flight_simulator.stan
https://github.com/stan-dev/example-models/blob/master/ARM/Ch.17/17.3_non-nested_models.R
```{r}
fit_program<-lmer(Program~1+(1|pair) +  (1|Judge),olympics_long)
```

```{r}
dataList.1 <- list(N=49, n_judges=7, n_pairs=7,  judge=as.integer(olympics_long$Judge), pair=as.integer(olympics_long$pair), y=olympics_long$Program)
                   
skating_stan<-"
data {
  int<lower=0> N;
  int<lower=0> n_judges;
  int<lower=0> n_pairs;
  int<lower=0,upper=n_judges> judge[N];
  int<lower=0,upper=n_pairs> pair[N];
  vector[N] y;
}
parameters {
  real<lower=0> sigma;
  real<lower=0> sigma_gamma;
  real<lower=0> sigma_delta;
  vector[n_judges] gamma;
  vector[n_pairs] delta;
  real mu;
}
model {
  vector[N] y_hat;

  sigma ~ uniform(0, 100);
  sigma_gamma ~ uniform(0, 100);
  sigma_delta ~ uniform(0, 100);

  mu ~ normal(0, 100);
  
  gamma ~ normal(0, sigma_gamma);
  delta ~ normal(0, sigma_delta);

  for (i in 1:N)
    y_hat[i] = mu + gamma[judge[i]] + delta[pair[i]];
  y ~ normal(y_hat, sigma);
}
"
```

pilots <- read.table ("http://www.stat.columbia.edu/~gelman/arm/examples/pilots/pilots.dat", header=TRUE)

flight_simulator.sf1 <- stan(   model_code=skating_stan
, data=dataList.1, iter=2000, chains=4)


##  Multilevel logistic regression 

The folder `speed.dating` contains data from an experiment on a few hundred students that randomly assigned each participant to 10 short dates with participants of the opposite sex (Fisman et al., 2006). For each date, each person recorded several subjective numerical ratings of the other person (attractiveness, compatibility, and some other characteristics) and also wrote down whether he or she would like to meet the other person again. Label $y_{ij} = 1$ if person $i$ is interested in seeing person $j$ again $0$ otherwise.
And $r_{ij1},\dots, r_{ij6}$ as person $i$'s numerical ratings of person $j$ on the dimensions of attractiveness, compatibility, and so forth.
Please look at 
http://www.stat.columbia.edu/~gelman/arm/examples/speed.dating/Speed%20Dating%20Data%20Key.doc
for details.

```{r}
dating<-fread("http://www.stat.columbia.edu/~gelman/arm/examples/speed.dating/Speed%20Dating%20Data.csv")
dating_pooled <- glm(match~attr_o +sinc_o +intel_o +fun_o +amb_o +shar_o,data=dating,family=binomial)
dating_pooled <- glmer(match~gender + attr_o +sinc_o +intel_o +fun_o +amb_o +shar_o+(1|iid)+(1|pid),data=dating,family=binomial)
```

1. Fit a classical logistic regression predicting $Pr(y_{ij} = 1)$ given person $i$'s 6 ratings of person $j$. Discuss the importance of attractiveness, compatibility, and so forth in this predictive model.

```{r}
log.1 <- glm(match~attr_o +sinc_o +intel_o +fun_o +amb_o +shar_o,data=dating,family=binomial)
summary(log.1)
coefplot(log.1)
```

The result can be written:

$$P(match=1)=logit^{-1}(-5.6+0.22attr-0.02sinc+0.07intel+0.25fun-0.12amb+0.21shar)$$.


2. Expand this model to allow varying intercepts for the persons making the evaluation; that is, some people are more likely than others to want to meet someone again. Discuss the fitted model.
```{r}
log.mix <- lmer(match~gender+scale(attr_o) +scale(sinc_o) +scale(intel_o) +scale(fun_o) +scale(amb_o) +scale(shar_o)+(1|iid),data=dating,family=binomial(link="logit"))
summary(log.mix)
```

3. Expand further to allow varying intercepts for the persons being rated. Discuss the fitted model.
```{r}
log.2 <- glmer(match~gender+scale(attr_o) +scale(sinc_o) +scale(intel_o) +scale(fun_o) +scale(amb_o) +scale(shar_o)+(1|iid)+(1|pid),data=dating,family=binomial)
summary(log.2)
a_iid <- data.frame(ranef(log.2))[1:5,4]
a_pid <- data.frame(ranef(log.2))[552:556,4]
a_iid
a_pid
```



4. You will now fit some models that allow the coefficients for attractiveness, compatibility, and the other attributes to vary by person. Fit a no-pooling model: for each person i, fit a logistic regression to the data $y_{ij}$ for the 10 persons j whom he or she rated, using as predictors the 6 ratings $r_{ij1},\dots,r_{ij6}$ . (Hint: with 10 data points and 6 predictors, this model is difficult to fit. You will need to simplify it in some way to get reasonable fits.)
```{r}
no_pooling <- glm(match~attr_o + sinc_o + intel_o + fun_o + amb_o + shar_o + factor(iid)-1,data=dating)
```


5. Fit a multilevel model, allowing the intercept and the coefficients for the 6 ratings to vary by the rater i.
```{r}

```

6. Compare the inferences from the multilevel model in (5) to the no-pooling model in (4) and the complete-pooling model from part (1) of the previous exercise.
```{r}
```

